---
title: "k-NN"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Problem:-
Universal bank is a young bank growing rapidly in terms of overall customer acquisition. The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite small, and the bank is interested in expanding this base  rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.

A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.

The file UniversalBank.csv contains data on 5000 customers. The data include customer demographic information (age, income, etc.), the customer’s relationship with the bank (mortgage, securities account, etc.), and the customer response to the last personal loan campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the personal loan that was offered to them in the earlier campaign.

Solution:-

*Using Libraries needed and reading the data :-

```{r}

library(caret)
library(dplyr)
library(FNN)

UniversalBank <- read.csv("UniversalBank.csv")
head(UniversalBank)
str(UniversalBank)

```

Q1) Consider the following customer:
Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?

* Transforming categorical predictors into dummy variables and Elementing the ID and ZIP code:-

```{r}

UniversalBank <- UniversalBank[,c(-1,-5)]
str(UniversalBank)

#Dummy Variables

test_0 <- UniversalBank$Education
Testfactor <- factor(test_0, labels=c("E1","E2","E3"))
str(Testfactor)
UniversalBank[["Education"]] <- Testfactor
levels(UniversalBank$Education)

Dummy<- dummyVars("~Education", data=UniversalBank)
y <- predict(Dummy,newdata=UniversalBank)
UniversalBank <- cbind(UniversalBank, y)
UniversalBank <- UniversalBank[, c(-6)]

```

* Partitioning the data into training as 60% and validation as 40% sets; Normalization of the data and k-NN classification using k=1 :-

```{r}

set.seed(10)
partition <- createDataPartition(UniversalBank$Personal.Loan, p=0.6, list=FALSE)
Train_Data <- UniversalBank[partition,]
dim(Train_Data)
Validation_data <- UniversalBank[-partition,]
dim(Validation_data)

#Normalization

Norm<-preProcess(Train_Data, method = c('center', 'scale'))
Train_Norm<-predict(Norm,Train_Data)
summary(Train_Norm)
Validation_Norm<-predict(Norm,Validation_data)
summary(Validation_data)

Train_Predictors<-Train_Norm[,-7]
validation_Predictors<-Validation_Norm[,-7]
Train_labels <-factor(Train_Data[,7], levels=c(0,1), labels = c("NO","YES"))
Validation_labels <- factor(Validation_data[,7], levels=c(0,1), labels = c("NO","YES"))

#K-NN classification using k=1

PredictedTest_labels <- knn(Train_Predictors,validation_Predictors,cl=Train_labels,k=1,prob = TRUE)
str(PredictedTest_labels)

#Finding particular values for variables-> 1=loan acceptance, 0=loan denial
test_1 <- c(40, 10, 84, 2, 2, 0, 0, 0, 1, 1, 0, 1, 0)
PredictedTest_label2 <- knn(Train_Predictors, test_1, cl=Train_labels, k=1, prob = TRUE)
str(PredictedTest_label2)

```

Q2) What is a choice of k that balances between overfitting and ignoring the predictor information?

* Finding hypertuned 'k' value to get optimal k-NN :-
```{r}

library(e1071)
accuracy.df <- data.frame(k = seq(1, 10, 1), accuracy = rep(0, 10))
Validation_labels<-as.factor(Validation_labels)

for(i in 1:10) {
  PredictedTest_labels3 <- knn(Train_Predictors, validation_Predictors, cl=Train_labels, k=i,prob = TRUE)
  accuracy.df[i, 2] <- confusionMatrix(PredictedTest_labels3, Validation_labels)$overall[1] 
}
accuracy.df
which.max(accuracy.df$accuracy)

```

Q3) Show the confusion matrix for the validation data that results from using the best k.

*Confusion matrix using hypertuned 'k' value i.e. k=3 :-

```{r}

library("gmodels")
PredictedTest_label4 <- knn(Train_Predictors, validation_Predictors, cl=Train_labels, k=3)
CrossTable(x=Validation_labels,y=PredictedTest_label4 ,prop.chisq = FALSE) 

```
#Accuracy = 96.8%

Q4)Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.


* Classified the customers using k=3 :- 
```{r}

test_2 <- c(40, 10, 84, 2, 2, 0, 0, 0, 1, 1, 0, 1, 0)
PredictedTest_label3 <- knn(Train_Predictors,test_2 , cl=Train_labels, k=3, prob = TRUE)
str(PredictedTest_label3)

```
