---
title: "R/Git"
author: "Amruta"
date: "10/28/2019"
output: html_document
---
Problem :-

This assignment will concentrate on using R and Git. Specifically, you will do the following:

1) Download a dataset from the web. You may use any source, but specify the source in       your code. Also ensure that the data has a mix of quantitative and qualitative           (categorical) variables.
2) Import the dataset into R
3) Print out descriptive statistics for a selection of quantitative and categorical         variables.
4) Transform at least one variable. It doesn't matter what the transformation is.
5) Plot at least one quantitative variable, and one scatterplot
6) Upload your R program, and any associated datafiles to your git account. Remember to     create    a separate repository for this class.
7) Paste the address to your repository in the answer box here in BB.

Solution :-

* Downloading a dataset from the web & Importing the dataset into R :- 

```{r}

#source-https://dasl.datadescription.com/datafiles/?_sf_s=archery&_sfm_cases=4+59943

Archery<-read.csv("Archery Data.csv") #assign variable

```

* Descriptive statistics for a selection of quantitative and categorical variables :-

```{r}

# Mean
mean(Archery$Total_score)

# Median
median(Archery$X1st_Half)

# Range
range(Archery$X1st_Half)

# Variance
var(Archery$X2nd_Half)

# Standard Deviation
sd(Archery$X2nd_Half)

# Highest of total score amongst all Archer Women
max(Archery$Total_score)

# Lowest of total score amongst X1st half
min(Archery$X1st_Half)

# Which archer has highest total score
which.max(Archery$Total_score)

# Which Archer women has highest total score(shows all the columns)
Archery[which.max(Archery$Total_score),]

# Which Archer women has lowest total score(shows all the columns)
Archery[which.min(Archery$Total_score),]

# To get index of the X1st Half (order always sort values ascendingly)
order(Archery$X1st_Half)

```

* Transformation of variables :-

```{r}

library(rcompanion)

#Log transformation

Total_log <-log(Archery$Total_score)
Total_log
plotNormalHistogram(Total_log, main = "Total Score")

#squareroot transformation

Total_sqrt = sqrt(Archery$Total_score)
Total_sqrt
plotNormalHistogram(Total_sqrt, main = "Total Score")

```

* Plotting the quantitative variable and a scatter plot :-

```{r}

# Bar plot

barplot(Archery$Total_score,xlab="Count",ylab = "Total score",main = "Count Vs Total score")

# Scatter plot

plot.default(Archery$X1st_Half,Archery$X2nd_Half, xlab = "X1st Half",ylab = "X2nd Half",main ="X1st Half Vs X2nd Half")

```

